---
title: "Covid91 vaccine study Final2022"
author: "Carmen Ramos-Caminos"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
People <-read.csv(file="final_Data.csv",header=TRUE)
Females = filter(FinalData,sex=="F")
Males = filter(FinalData,sex=="M")
LGBTQ = filter(FinalData,LGBTQ=="straight")
LGBTQ = filter(FinalData,LGBTQ=="gay")
DrugUser = filter(FinalData, DrugUser=="no")
DrugUser = filter(FinalData, DrugUser=="yes")
Infected = filter(FinalData, infected=="covid19")
```


# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.  

Is the new preventative vaccine (covid91) treatment effective for the following subgroups sex (male/female), LGBTQ, DrugUser? 

$H_0$ Preventative covid91 vaccine treatment has no effect on subgroup: sex (Male/Female): LGBTQ: DrugUser 

$H_a$ Preventative covid91 treatment affects subgroup: sex (male/female): LGBTQ: DrugUser 

# Methods 

Both treatment and subgroups sex(Male/Female), LGBTQ, DrugUser are categorical variables. Therefore, the method of analysis technique used is cat~cat. I will use a barchart for graphical results , a crosstable for numerical results, a chi-square test and fisher exact test for inference.  

# Descriptive Results

## Males   

### Graphic Descriptive Results   

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```  
  
The bar chart shows that Males who take the vaccine (covid91) are far less likely to contract the virus covid19. Males who take the placebo are 68% more likely to contract covid19 versus males who took the vaccine and contacted covid19 32% of the time. 

### Numerical Descriptive Results  

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```  
  
The percentage of infection is much higher for Males who took the placebo. Those who took the placebo became infected approximately 67.78%  of the time versus 32.22% who took the vaccine and contracted covid19. The Males in the placebo group are 2 times more likely to become infected when receiving the placebo.
  
### Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```  

I reject the $H_0$ Null Hypothesis, and accept the $H_a$ Alternative hypothesis as true. There is a significant negative effect when Males took the placebo vs the covid91 vaccine which has a positive effect.  The Chi-squared test shows that the P-value is 0.00000000004812 an amount significantly less than .05. The Fisher exact test shows the P-Value is 0.00000000002645 also an number significalty less than 0.05. The test also confirms that the odds ratio of Males who take the placebo contract covid19 2 to 1 more times. The covid91 vaccine is effective for the Male subgroup tested. I would recommend the covid91 vaccine for the Male subgroup.   

## Efficacy

```{r}
efficMales = 2.09/(2.09 + 1) * 100
efficMales = round(efficMales)

confIntLower = 1.67/(1.67 + 1) * 100
confIntLower = round(confIntLower)
confIntUpper = 2.64/(2.64 + 1) * 100
confIntUpper = round(confIntUpper)
```

The efficacy for Males was `r efficMales`%

The 95% confidence interval for efficacy was (`r confIntLower` , `r confIntUpper`)%


## Females 

### Graphic Descriptive Results 

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```  

The bar chart shows that Females who take the vaccine (covid91) are less likely to contract the virus covid19. Females who took the placebo are 58% more likely to contract covid19 versus Females who took the vaccine and contacted covid19 42% of the time.

### Numerical Descriptive Results 

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```  

The percentage of infection is much higher for Females who took the placebo. Those who took the placebo became infected approximately 58.89%  of the time versus 41.11% who took the vaccine and contracted covid19. The Females in the placebo group are 1.5 times more likely to become infected when receiving the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```  

I reject the $H_0$ Null Hypothesis, and accept the $H_a$ Alternative hypothesis as true. There is a  negative effect when Females took the placebo vs the covid91 vaccine which has a positive effect.  The Chi-squared test shows that the P-value is 0.0011 an amount less than .05. The Fisher exact test shows the P-Value is 0.0009484 also an number less than 0.05. The test also confirms that the odds ratio of females who take the placebo contract covid19 almost 1.4 more times than Females who took the vaccine. The covid91 vaccine is effective for the Female subgroup tested. I would recommend the Covid91 for the Female subgroup. 

## Efficacy 

```{r}
efficFemales = 1.42/(1.42 + 1) * 100
efficFemales = round(efficFemales)

confIntLower = 1.15/(1.15 + 1) * 100
confIntLower = round(confIntLower)
confIntUpper = 1.77/(1.77 + 1) * 100
confIntUpper = round(confIntUpper)
```

The efficacy for Females was `r efficFemales`%

The 95% confidence interval for efficacy was (`r confIntLower` , `r confIntUpper`)%


## LGBTQ 

### Graphic Descriptive Results  

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```  

The bar chart shows that in the individuals in subgroup LGBTQ  who take the vaccine (covid91) are more likely to contract the virus covid19. Subgroup LGBTQ who take the placebo are 13% less likely to contract covid19 versus LGBTQ subgroup who took the vaccine and contacted covid19 87% of the time. 


### Numerical Descriptive Results 

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```  

The percentage of infection is higher for subgroup LGBTQ who took the vaccine. Those who took the placebo became infected approximately 13.33%  of the time versus 86.67% who took the vaccine and contracted covid19. The subgroup LBGTQ in the vaccine group are 0.1465 times more likely to become infected when receiving the vaccine.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```    

I reject the $H_0$ Null Hypothesis. Preventative vaccine, covid91, treatment has a negative effective for subgroup LGBTQ. Members of the LGBTQ subgroup were more likely to contract covid 19 after taking the vaccine. The P-Value for Chi-Sqaured test P-Value is 0.0000000132 less than 0.05. the Fisher exact test P-value is 0.000000001642. The odds ratio is 0.1465. I would not recommend that memebers of the LGBTQ subgroup take the covid91 vaccine.  

## Efficacy 

```{r}
efficLGBTQ = 0.14/(0.14 + 1) * 100
efficLGBTQ = round(efficLGBTQ)

confIntLower = 0.05/(0.05 + 1) * 100
confIntLower = round(confIntLower)
confIntUpper = 0.31/(0.31 + 1) * 100
confIntUpper = round(confIntUpper)
```

The efficacy for LGBTQ was `r efficLGBTQ`%

The 95% confidence interval for efficacy was (`r confIntLower` , `r confIntUpper`)%


## Druggies

### Graphic Descriptive Results  

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
``` 

The bar chart shows in subgroup DrugUser individuals who take the vaccine (covid91) are more less likely to contract the virus covid19. DrugUsers who take the vaccine are 85% more likely to contract covid19 versus DrugUsers who took the placebo and contacted covid19 15% of the time. 


### Numerical Descriptive Results 

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```        

The percentage of infection is higher for DrugUsers who took the vaccine. Those who took the placebo became infected approximately 15%  of the time versus 85% who took the vaccine and contracted covid19. The DrugUsers in the placebo group are  less likely to become infected when receiving the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```     

## Efficacy

```{r}
efficDrugUser = .17/(.17 + 1) * 100
efficDrugUser = round(efficDrugUser)

confIntLower = .07/(.07 + 1) * 100
confIntLower = round(confIntLower)
confIntUpper = .36/(.36 + 1) * 100
confIntUpper = round(confIntUpper)
```

The efficacy for DrugUser was `r efficDrugUser`%

The 95% confidence interval for efficacy was (`r confIntLower` , `r confIntUpper`)%


I reject the $H_0$ Null Hypothesis. Preventative vaccine, covid91, treatment has a negative effective for subgroup DrugUser. Members of the DrugUser subgroup were more likely to contract covid 19 after taking the vaccine. The P-Value for Chi-Squared test P-Value is 0.0000001803 less than 0.05. the Fisher exact test P-value is 0.00000002886. The odds ratio is 0.1775. I would not recommend that members of the DrugUser subgroup take the covid91 vaccine. 

# Discussion and Conclusion 

The question posed asked to determine whether the covid91 vaccine affected the subgroups Male, Female, LGBTQ, and DrugUser tested. Additionally, we are asked to recommend or not recommend the covid91 vaccine for the four subgroups. The Male subgroup was affected positively by the vaccine. Males were less likely to contract covid after receiving the vaccine versus the placebo. The Female subgroup showed a positive effect. Females who took the vaccine were also affected positively compared to those who took the placebo. Therefore, I conclude that both Males and Females should take the covid vaccine. 

The next subgroup includes the LGBTQ subgroup who were affected negatively by the covid vaccine. LGBTQ individuals were more likely to contract covid after taking the vaccine. The placebo seemed more effective for this subgroup. The subgroup DrugUser also showed a negative affect when they took the covid vaccine versus the placebo. Members of this group were more likely to contract covid after taking the the vaccine, Therefore, I conclude that members of the LGBTQ subgroup and members of the DrugUser subgroup should not take the vaccine as they are more likey to contract covid.

The vaccine had an affect on each of the subgroups. Male and Female were affected positively. I would recommend that these two groups take the covid91 vaccine. The LGBTQ and DrugUser  subgroup were affected negatively. I don not recommend that either of these two subgroups take the covid91 vaccine. The results showed that there is a difference in each of these subgroups. Therefore in each case I rejected the Null Hypothesis.  












